---
title: "Plots"
author: "A. Warren, Jimmy Hou"
date: "2020/12/3"
output: html_document
---
```{r load-packages, echo = FALSE, messages = FALSE, results = FALSE, warning = FALSE}
library(tidyverse)
library(tidymodels)
```

```{r load-data, echo = FALSE, messages = FALSE, results = FALSE, warning = FALSE}
teams_cleaned <- read_csv("/cloud/project/data/teams-cleaned.csv")

#doing training split
set.seed(3)
teams_split <- initial_split(teams_cleaned, prop = 0.80)
teams_train <- training(teams_split)
teams_test <- testing(teams_split)

#have a lot of NAs for drs and uzr/uzr150/def, so he's something to deal with
#that:
teams_train %>%
  filter(is.na(mean_drs) == FALSE) -> teams_train_drs
teams_train %>%
  filter(is.na(mean_uzr) == FALSE) -> teams_train_uzr
```

### Initial Plots vs Win Rate

With linear models overlayed and residual vs fitted plots. For the time being,
only considering single-variable regression.

##### Batting Average (AVG)

```{r batting-average, echo = FALSE}
#It's the same process for every variable, so I'll just walk
#through the first one.

#create graph, add linear model trendline
teams_train %>%
  ggplot(mapping = aes(x = mean_avg , y = win_rate)) +
  geom_point(size = 1.2, alpha = 0.35) +
  geom_smooth(method = "lm")

#linear model, note R squared and adjusted R squared
avg_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(win_rate ~ mean_avg, data = teams_train)
tidy(avg_fit)
glance(avg_fit)

#residuals analysis
avg_aug <- augment(avg_fit$fit)
ggplot(data = avg_aug,
       aes(x = .fitted,
           y = .resid)) +
  geom_jitter() +
  geom_hline(yintercept = 0,
             linetype = "dashed")
```

##### Slugging Percentage (SLG)

```{r slugging-percentage, echo = FALSE}
teams_train %>%
  ggplot(mapping = aes(x = mean_slg , y = win_rate)) +
  geom_point(size = 1.2, alpha = 0.35) +
  geom_smooth(method = "lm")

slg_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(win_rate ~ mean_slg, data = teams_train)
tidy(slg_fit)
glance(slg_fit)

slg_aug <- augment(slg_fit$fit)
ggplot(data = slg_aug,
       aes(x = .fitted,
           y = .resid)) +
  geom_jitter() +
  geom_hline(yintercept = 0,
             linetype = "dashed")
```

##### Isolated Power (ISO)

```{r isolated-power, echo = FALSE}
teams_train %>%
  ggplot(mapping = aes(x = mean_iso , y = win_rate)) +
  geom_point(size = 1.2, alpha = 0.35) +
  geom_smooth(method = "lm")

iso_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(win_rate ~ mean_iso, data = teams_train)
tidy(iso_fit)
glance(iso_fit)

iso_aug <- augment(iso_fit$fit)
ggplot(data = iso_aug,
       aes(x = .fitted,
           y = .resid)) +
  geom_jitter() +
  geom_hline(yintercept = 0,
             linetype = "dashed")
```

##### Batting Average on Balls in Play (BABIP)

```{r batting-average-on-balls-in-play, echo = FALSE}
teams_train %>%
  ggplot(mapping = aes(x = mean_babip , y = win_rate)) +
  geom_point(size = 1.2, alpha = 0.35) +
  geom_smooth(method = "lm")

babip_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(win_rate ~ mean_babip, data = teams_train)
tidy(babip_fit)
glance(babip_fit)

babip_aug <- augment(babip_fit$fit)
ggplot(data = babip_aug,
       aes(x = .fitted,
           y = .resid)) +
  geom_jitter() +
  geom_hline(yintercept = 0,
             linetype = "dashed")
```

##### On Base Percentage (OBP)

```{r on-base-percentage, echo = FALSE}
teams_train %>%
  ggplot(mapping = aes(x = mean_obp , y = win_rate)) +
  geom_point(size = 1.2, alpha = 0.35) +
  geom_smooth(method = "lm")

obp_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(win_rate ~ mean_obp, data = teams_train)
tidy(obp_fit)
glance(obp_fit)

obp_aug <- augment(obp_fit$fit)
ggplot(data = obp_aug,
       aes(x = .fitted,
           y = .resid)) +
  geom_jitter() +
  geom_hline(yintercept = 0,
             linetype = "dashed")
```

##### Weighted On Base Average (wOBA)

```{r weighted-on-base-average, echo = FALSE}
teams_train %>%
  ggplot(mapping = aes(x = mean_w_oba , y = win_rate)) +
  geom_point(size = 1.2, alpha = 0.35) +
  geom_smooth(method = "lm")

woba_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(win_rate ~ mean_w_oba, data = teams_train)
tidy(woba_fit)
glance(woba_fit)

woba_aug <- augment(woba_fit$fit)
ggplot(data = woba_aug,
       aes(x = .fitted,
           y = .resid)) +
  geom_jitter() +
  geom_hline(yintercept = 0,
             linetype = "dashed")
```

##### Weighted Runs Created (wRC)

```{r weighted-runs-created, echo = FALSE}
teams_train %>%
  ggplot(mapping = aes(x = mean_w_rc , y = win_rate)) +
  geom_point(size = 1.2, alpha = 0.35) +
  geom_smooth(method = "lm")

wrc_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(win_rate ~ mean_w_rc, data = teams_train)
tidy(wrc_fit)
glance(wrc_fit)

wrc_aug <- augment(wrc_fit$fit)
ggplot(data = wrc_aug,
       aes(x = .fitted,
           y = .resid)) +
  geom_jitter() +
  geom_hline(yintercept = 0,
             linetype = "dashed")
```

##### Base Runs (BsR)

```{r base-runs, echo = FALSE}
teams_train %>%
  ggplot(mapping = aes(x = mean_bsr , y = win_rate)) +
  geom_point(size = 1.2, alpha = 0.35) +
  geom_smooth(method = "lm")

bsr_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(win_rate ~ mean_bsr, data = teams_train)
tidy(bsr_fit)
glance(bsr_fit)

bsr_aug <- augment(bsr_fit$fit)
ggplot(data = bsr_aug,
       aes(x = .fitted,
           y = .resid)) +
  geom_jitter() +
  geom_hline(yintercept = 0,
             linetype = "dashed")
```

##### Offense (Off)

```{r offense, echo = FALSE}
teams_train %>%
  ggplot(mapping = aes(x = mean_off , y = win_rate)) +
  geom_point(size = 1.2, alpha = 0.35) +
  geom_smooth(method = "lm")

off_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(win_rate ~ mean_off, data = teams_train)
tidy(off_fit)
glance(off_fit)

off_aug <- augment(off_fit$fit)
ggplot(data = off_aug,
       aes(x = .fitted,
           y = .resid)) +
  geom_jitter() +
  geom_hline(yintercept = 0,
             linetype = "dashed")
```

##### Defensive Runs Saved (DRS)

```{r defensive-runs-saved, echo = FALSE}
teams_train_drs %>%
  ggplot(mapping = aes(x = mean_drs , y = win_rate)) +
  geom_point(size = 1.2, alpha = 0.35) +
  geom_smooth(method = "lm")

drs_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(win_rate ~ mean_drs, data = teams_train_drs)
tidy(drs_fit)
glance(drs_fit)

drs_aug <- augment(drs_fit$fit)
ggplot(data = drs_aug,
       aes(x = .fitted,
           y = .resid)) +
  geom_jitter() +
  geom_hline(yintercept = 0,
             linetype = "dashed")
```

##### Ultimate Zone Rating (UZR)

```{r ultimate-zone-rating, echo = FALSE}
teams_train_uzr %>%
  ggplot(mapping = aes(x = mean_uzr , y = win_rate)) +
  geom_point(size = 1.2, alpha = 0.35) +
  geom_smooth(method = "lm")

uzr_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(win_rate ~ mean_uzr, data = teams_train_uzr)
tidy(uzr_fit)
glance(uzr_fit)

uzr_aug <- augment(uzr_fit$fit)
ggplot(data = uzr_aug,
       aes(x = .fitted,
           y = .resid)) +
  geom_jitter() +
  geom_hline(yintercept = 0,
             linetype = "dashed")
```

##### Ultimate Zone Rating per 150 Games (UZR/150)

```{r uzr-150, echo = FALSE}
teams_train_uzr %>%
  ggplot(mapping = aes(x = mean_uzr150 , y = win_rate)) +
  geom_point(size = 1.2, alpha = 0.35) +
  geom_smooth(method = "lm")

uzr150_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(win_rate ~ mean_uzr150, data = teams_train_uzr)
tidy(uzr150_fit)
glance(uzr150_fit)

uzr150_aug <- augment(uzr150_fit$fit)
ggplot(data = uzr150_aug,
       aes(x = .fitted,
           y = .resid)) +
  geom_jitter() +
  geom_hline(yintercept = 0,
             linetype = "dashed")
```

##### Defense (Def)

```{r defense, echo = FALSE}
teams_train_uzr %>%
  ggplot(mapping = aes(x = mean_def , y = win_rate)) +
  geom_point(size = 1.2, alpha = 0.35) +
  geom_smooth(method = "lm")

def_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(win_rate ~ mean_def, data = teams_train_uzr)
tidy(def_fit)
glance(def_fit)

def_aug <- augment(def_fit$fit)
ggplot(data = def_aug,
       aes(x = .fitted,
           y = .resid)) +
  geom_jitter() +
  geom_hline(yintercept = 0,
             linetype = "dashed")
```

##### Wins Above Replacement (WAR)

```{r wins-above-replacement, echo = FALSE}
teams_train %>%
  ggplot(mapping = aes(x = mean_war , y = win_rate)) +
  geom_point(size = 1.2, alpha = 0.35) +
  geom_smooth(method = "lm")

war_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(win_rate ~ mean_war, data = teams_train)
tidy(war_fit)
glance(war_fit)

war_aug <- augment(war_fit$fit)
ggplot(data = war_aug,
       aes(x = .fitted,
           y = .resid)) +
  geom_jitter() +
  geom_hline(yintercept = 0,
             linetype = "dashed")
```


### Analysis of Initial Plots

Observation of residual vs. fitted plots shows no discernible patterns,
suggesting that the linear model is an adequate fit in all cases. Observation of
adjusted R squared values shows that the linear models of WAR, Off, and wRC vs
Win Rate are best able to account for variance in data (values of 0.534, 0.442,
and 0.445 respectively). Observation of plots does not suggest any cases of low
R squared resulting from a high number of outliers.

### Additional Plots

Examining how well the models for WAR, wRC, and Off function as predictors.

```{r wins-above-replacement-prediction, echo = FALSE}
war_pred <- predict(war_fit, teams_test) %>%
  bind_cols(teams_test %>% select(win_rate, mean_war))
tibble(war_pred)

#Not quite sure how to evaluate the accuracy of these predictions, so I'm going 
#to look at the difference between predicted and actual, then calculate key 
#summary statistics.
war_pred %>%
  mutate(pred_diff = abs(.pred - win_rate)) -> war_pred2
tibble(war_pred2)
mean(war_pred2$pred_diff)
median(war_pred2$pred_diff)
IQR(war_pred2$pred_diff)
sd(war_pred2$pred_diff)
```




