---
title: "Project proposal"
author: "tut2_team5"
output: github_document
---

```{r load-packages, message = FALSE, echo = FALSE}
library(tidyverse)
library(broom)
library(skimr)
```

```{r load-data, message = FALSE, echo = FALSE}
people <- read_csv("/cloud/project/data/baseballdatabank-master/core/People.csv")

batting <- read_csv("/cloud/project/data/baseballdatabank-master/core/Batting.csv",
                    col_types = cols(HBP = col_integer(), 
                                     SH  = col_integer(),
                                     SF  = col_integer(),
                                     IBB  = col_integer(),
                                     .default = "?"))

# Make sure you choose the right type before loading a dataset.
# read_csv() determines the type from the first 1000 entries.
# e.g. HBP is usually 0 or 1 for most so it thinks it is a logical.
# However, later in the dataset it sees a 2 so it throws a parsing error.
```

## 1. Introduction

*"I made baseball as much fun as doing your taxes."* 
- Bill James, father of modern Sabermetrics (The Simpsons S22E03 "MoneyBART")

Statistical study of baseball (also known as Sabermetrics or SABRmetrics) has become an increasingly popular tool both for fans seeking to understand the game and for organizations seeking to optimize their rosters and build better teams. Our group will look at how effectively different measurements for evaluating individual player performance predict team success.

The datasets we are going to use were compiled by an organization Chadwick Baseball Bureau. From their [webpage](http://chadwick-bureau.com/): "Chadwick Baseball Bureau was founded to support baseball professionals and analysts by providing tidy baseball data. We create value for our clients and for the community by allowing users to focus on understanding and acting on information, rather than spending their time compiling and integrating it." The CBB dataset includes data from [Retrosheet](http://retrosheet.org/), another organization dedicated to providing baseball data. Much of this compiled data was originally collected from MLB teams and the league itself, with other sources such as sportswriters contributing to game data. 

In the course of this project we won't cover data from before the league and teams started keeping computer records, so the data we cover will entirely originate from these official records.

## 2. Data

Information regarding the licensing of this data can be found in ```/project/data/baseballdatabank-master/README.txt```

The dataset is large and is composed of 29 csv files. For convenience, I'll only include a skim of player data joined with batting data. 

```{r, message = FALSE, echo = FALSE}
batting_demo <- full_join(people, batting)
skim(batting_demo)
```


## 3. Data analysis plan

We will look at the effectiveness of different player stats in predicting team success. To begin with, the dataset only includes basic statistics. Most of the stats we're interested in (OBP, OPS, wOBA, wRC, wRAA, etc.) will have to be calculated from these basic statistics. In order to simplify the data, we'll use player rosters to find mean value and other summary statistics for each variable for each team over time. Plotting each team's mean value for each variable against game outcomes to gauge strength of association should provide a starting point for assessing the effectiveness of each variable in predicting game outcomes. We can follow up by looking for possible confounding variables and controlling for them (e.g. possible impact of morale on team performance could be gauged by looking into win/loss streaks and comparing home vs. away game performance).

If we've got extra time we could run the process for a few different seasons. Conversely, if it's too much work we might consider cutting defensive stats since they seem like they might be a hassle to calculate.
